Spectroscopic confirmation of very-wide binaries and large-separation comoving stars from \textsl{Gaia DR1}

### 2017-07-11 paper/comoving-followup.tex
Add smoh comments

@@

\begin{abstract}
% Context
Widely-separated, coeval stars are powerful tools for testing star-formation
model predictions, improving stellar atmosphere models, and inferring the
mass distribution of the Milky Way.
They vastly outnumber the Galactic open clusters, but are still expected to
share a common origin and chemistry.
The kinematic properties of bound wide binaries and their unbound complements
serve as potentiometers for the smooth tidal field and perturbation spectrum in
the Milky Way.
Thus far, a few hundred genuine comoving star pairs with separations as large
as $8~\pc$ have been discovered in the solar neighborhood.
% Aims
In previous work, we have identified $\sim 10,000$ candidate comoving star pairs
using only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS)
catalog, a part of \gaia\ \DR{1}.
Without radial velocity measurements for the vast majority of the pairs, the
contamination rate at large separations is significant ($\sim 50\%$).
% Methods
Here, we present results from our own radial-velocity survey of a sample of the
candidate pairs to validate wide-separation comoving star pairs.
% Results
Of the 311 observed comoving pairs, we identify 128 genuine comoving pairs with
separations as large as $\approx 10~\pc$.
% Conclusions
We find that the number of comoving star pairs with very wide separations
appears flat in linear separation at separations $\gtrsim 1~\pc$, strongly
suggesting the existence of a population of recently unbound wide binaries or
stellar associations.
% Gaia DR2 -> how many do we expect? Can we say anything? Out to what distance and
% separation for a typical MS star can we determine they are bound assuming
% high-res spectra to get very precise RVs?
\end{abstract}

@@

The majority of stars are thought to form in dynamically cold, chemically
homogeneous clusters, at least relative to the kinematic and chemical
dispersions observed amongst typical stars in the Galactic disk.
Because star formation proceeds rapidly (e.g., \citealt{Bovy:2016}) and the
progenitor gas cloud disperses quickly (cite), it is typically assumed that
stars formed in the same birth cluster have the same initial chemistry
(e.g., cite + \citealt{Feng:2014}) and same ages.
These birth clusters are typically born un- or loosely-bound and therefore
quickly succumb to the tidal field of the Milky Way.
Obvious dynamical associations between stars from the same cluster are then
erased as the constituent stars mix into the general population of disk stars.
However, the stars do preserve their initial associations in age, chemical
abundance patterns, and dynamical invariants (e.g., orbital actions).
It has therefore been long presumed that ``chemical tagging'' and chemodynamical
models of the Galactic disk will allow for inferences about the formation and
evolution of the Milky Way (e.g., \citealt{todo}).
That is, the chemical environment of a birth cluster is likely imprinted on its
stellar children and preserved so that their initial association may be inferred
with precisely measured abundance patterns (of $\gtrsim 10$ species).

The feasibility of chemical tagging depends on the intrinsic chemical
dispersions of stars formed from the same birth clusters, the precision of the
abundance measurements, and the ability to tell that the chemical abundances of
two stars are more similar than two randomly-paired Galactic disk stars.
Recent work suggests that, at least for red giant stars, the abundance
distributions of 20 elements in a handful of open clusters included in the
\acronym{APOGEE} survey (\citealt{Majewski:2016,Nidever:2015}) are consistent
with having intrinsic dispersions smaller than the abundance measurement
precisions ($\approx 0.03~\dex$; \citealt{Bovy:2016,Ness:2017}).
In effect, for most elements, the intrinsic dispersions are unconstrained
because of measurement noise, so the true intrinsic dispersions are not known.
This demonstration of chemical uniformity in open clusters is promising for
chemical tagging.
However, it was also recently shown that $\approx 0.3\%$ of random pairings of
Galactic field red giant stars have abundance patterns that are
indistinguishable at the same level (``doppelgangers''), even with 20 chemical
species (\citealt{Ness:2017}).

Are these properties of open clusters universal, or limited to the few clusters
presently studied?
Do the chemical properties of open clusters depend on the star formation rate or
efficiency, which may evolve with time?
Ideal tests of the chemical homogeneity assumptions would use many more open
clusters with varied cluster masses and member stars that span a large range in
stellar masses and evolutionary stages.
Precise kinematic and color data from the \gaia\ mission
(\citealt{Gaia-Collaboration:2016,Gaia-Collaboration:2016a}) will help refine
existing membership catalogs and will lead to discoveries of many more open
clusters and young star associations (see, e.g., \citealt{Oh:2017}).
However, because they disrupt quickly, the present-day open cluster population
is biased towards younger or more massive clusters.

@@

In previous work, we have identified a sample of high-confidence, candidate
comoving star pairs (\citealt{Oh:2017}) using only astrometric information from
the \tgas\ catalog (\citealt{Michalik:2015,Gaia-Collaboration:2016a}).
These comoving pairs were selected using a cut on a marginalized likelihood
ratio computed for all nearly co-spatial pairs of stars.
Briefly, likelihood 1, $L_1$, considers a model in which the full-space
velocities of each star in the pair are identical and sampled from a disk-like,
Sun-relative velocity distribution, and likelihood 2, $L_2$, considers a model
in which the full-space velocities of each star were drawn independently from
the same velocity distribution used above. The computed likelihood ratio is
actually the the ratio of fully marginalized likelihoods,
$\mathcal{L}_1/\mathcal{L}_2$, after marginalizing over true velocity and
distance for each star in a given pair.
After a parallax signal-to-noise ratio (SNR) cut ($\pi/\sigma_\pi > 8$), we
compute the likelihood ratio for all stars within $10~\pc$ of each target star
from the \tgas\ catalog.
The likelihoods appropriately take into account the reported covariances and
uncertainties between the astrometric measurements in the \tgas\ catalog and
naturally handle geometric or projection effects in comparing spherical
velocity components for star pairs separated by large angles on the sky.
After applying a conservative cut on the likelihood ratio motivated by computing
the same likelihood ratio for random pairings of stars ($\ln
\mathcal{L}_1/\mathcal{L}_2 > 6$, we find 13,058 comoving star pairs with
physical separations between $10^3~\au$ and $10~\pc$.
Of these pairs, there are 10,606 unique sources, implying that a significant
fraction of the identified pairs are members of larger groups.
$\approx$4,000 of the pairs are mutually exclusive, i.e. neither member is
paired with other stars.
For a full description of the selection method, see \citealt{Oh:2017}.

@@

Spectra were obtained using the \project{Modspec} spectrograph mounted on the
$2.4~{\rm m}$ Hiltner telescope at MDM
observatory\footnote{\url{http://mdm.kpno.noao.edu}} on Kitt Peak (Arizona).
\project{Modspec} was set up in long-slit mode with a 1'' slit, a $600~{\rm
line}~{\rm mm}^{-1}$ grating, and a $2048\times2048~{\rm pixel}$ SITe CCD
detector (``Echelle'').
Only the central 300 pixel columns were read out from the CCD, and pixel rows
$>1600$ (at the blue end) were later trimmed because of the poor quantum
efficiency of the CCD at wavelengths $\lesssim 3600~{\rm \AA}$.
The resolution and wavelength range were $\approx2~{\rm \AA}~{\rm pixel}^{-1}$
from $\approx 3600$--$7200~{\rm \AA}$.
At this resolution, to obtain high signal-to-noise spectra, most exposures were
between $30$--$120~{\rm s}$ (15th and 85th percentiles).
To maximize the number of observed pairs, we therefore chose to not take
comparison lamp spectra at each pointing, which adds an overhead of $\approx
90~{\rm s}$ per observation.
To determine rough nightly wavelength solutions, spectra were obtained using
Hg-Ne and Ne comparison lamps at the beginning and end of each night, and
atmospheric emission lines were used to correct the wavelength solution on a
per-object basis.

We observed a total of 765 unique sources over 5 nights from (UT) 2017-03-11 to
2017-03-15, corresponding to 323 unique star pairs and $\approx 120$
calibration targets.
The calibration targets have previously measured radial velocities and chemical
abundances, which will be used for validation of our reduction pipeline and in
future work to measure metallicities and stellar parameters for each observed
source using \project{The Cannon} (\citealt{Ness:2015}).
The spectra were reduced and calibrated using a custom, publicly-available
pipeline written in \python, described below.

For each source or comparison lamp observation, the two-dimensional image is
bias corrected, flat-field corrected, and trimmed using standard routines
implemented in \package{ccdproc} (\citealt{Craig:2015}).
For comparison lamp observations, the central 100 pixel columns are extracted
and summed along the spatial axis, weighted by the inverse-variances of each
pixel.
Source spectral traces were always placed in this region during observations,
and the curvature of the wavelength solution over this region is negligible.
For source observations, the 2D spectra are extracted to 1D using an empirical
estimate of the pixel-convolved line spread function (LSF): at each of 16 pixel
rows between indices 750 and 850 (i.e. around the effective center of the
dispersion axis), a model for the LSF is fit to the 1D trace.
The LSF is assumed to be a Voigt profile with parameters $A$, amplitude; $x_0$,
profile centroid; $\sigma_G$, Gaussian root-variance; $\gamma_L$, Lorentzian
half-width at half maximum, and the background is modeled using a linear
polynomial with parameters $\alpha_1$ and $\alpha_2$, the slope and intercept
at the profile centroid, respectively.
The maximum-likelihood LSF + background model parameters are estimated from
each pixel row using \lbfgsb\ optimization (\citealt{Zhu:1994}), implemented in
the \package{scipy} package.
From the 16 fits, the median profile width parameters, $\sigma_G$ and
$\gamma_L$, are then taken to be the LSF width parameters.
At each of 1600 rows in each source spectrum image, the LSF amplitude and
centroid, and background slope and intercept at line centroid, are then fit
using the same procedure as above but with the width parameters fixed.
The LSF amplitudes are stored as the source fluxes, and the background model
intercepts are stored as the background (sky) fluxes.
Occasionally, nearby sources fall in the slit; these are masked by hand in the
2D images and do not bias the LSF extraction fits.

@@

This wavelength calibration procedure produces a nightly model for the
nonlinear dispersion of wavelength as a function of pixel value for each
extracted 1D spectrum.
However, because of flexure in the spectrograph, each observation will have an
additional offset of, typically, $\approx 0.5$--$2~{\rm \AA}$, corresponding to
a significant velocity offset at \Ha\ of $\approx 25$--$100~\kms$ (discussed
below).
\figurename~\ref{fig:sample-spectra} shows a random sample of 10
wavelength-calibrated source spectra of comoving star pair members normalized
to 1 at $\lambda = 5500~{\rm \AA}$ and shifted by index for visualization.
As suggested by the features in these 10 spectra, there is a mix of
metallicities and spectral types amongst the observed targets.
Note that the spectra are not flux-calibrated, as we only care about measuring
radial velocities from this sample.

@@

Of the 323 observed star pairs, 12 have at least one member with an unreliable
(flagged) radial velocity measurement from the pipeline above.
For the remaining 311 pairs, we use the measured radial velocities to re-compute
the marginal likelihood ratio for the two hypotheses considered in previous work
(see Appendix~B, \citealt{Oh:2017}).
We compute two fully marginalized likelihoods (FMLs) for the data: (1) assuming
that the two stars are comoving with identical 3D velocity vectors (within some
allowed small dispersion), and (2) assuming that the two stars have independent
3D velocity vectors drawn from a prior velocity distribution.
We set the small allowed dispersion ($s$, in previous work) to $1~\kms$, and use
a mixture of three Gaussian velocity components for the prior, meant to
represent the thin disk, thick disk, and halo velocity distributions relative to
the sun.

@@

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=0.7\linewidth]{llr-vs-dv.pdf}
  \end{center}
  \caption{%
    Point estimate of the 3D velocity difference between two stars in a comoving
    pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against the change in
    log-likelihood-ratio computed when including radial velocity data.
    Error bars are shown for a few randomly-chosen, representative pairs and
    represent the 15th and 85th percentiles of velocity differences computed
    from samples from the error distributions of parallax, proper motion
    components, and radial velocity.
    The likelihood ratio increases (i.e. the belief that the stars are comoving
    becomes stronger) when the 3D velocity difference is small.
    \label{fig:llr}}
\end{figure}

If the new likelihood ratio is larger than that computed from astrometry alone,
the added radial velocity information supports the belief that the two stars are
truly comoving in 3-space.
Conversely, if the new likelihood ratio is smaller, the radial velocity data
indicate that the pair is likely a false-positive and are not truly comoving.
In practice, we compute the logarithm of the likelihood ratios,
$\mathcal{R} = \ln \mathcal{L}_1/\mathcal{L}_2$; we refer to the
proper-motion-only log-FML-ratio for a given pair of stars as
\llrold\ (as computed in \citealt{Oh:2017})m and the log-FML-ratio including the radial velocity data as \llrnew.
We consider a pair as being genuinely comoving if $\llrnew > \llrold$---128 of
the 311 comoving pairs satisfy this constraint.
\figurename~\ref{fig:llr} shows the magnitude of the 3D velocity difference
plotted against the change in log-likelihood-ratio, $\llrnew - \llrold$, for all
observed pairs.
In detail, for visualization, we sample from the error distributions over proper
motion, parallax, and radial velocity for each star individually and compute the
magnitude of the 3D velocity differences, $|\bs{v}_1-\bs{v}_2|$, for $2^{16}$
samples.
We plot the median of the resulting distribution of velocity-differences as
dark markers (circles).
The error bars show the 15th and 85th percentiles of the distributions for a few
randomly chosen pairs.
Note that pairs with large 3D velocity differences have large (negative) changes
to their log-likelihood-ratio values.

@@

At separations $\gtrsim 1~\pc$, $\approx 40\%$ of the comoving pairs are
genuinely comoving; interestingly, the separation distribution of the genuine
pairs appears to be uniform or possibly growing in number with increasing
separation.
This strongly suggests the existence of a population of unbound wide binaries or
dissolved open clusters, which maintain approximate spatial coherence for long
times because of the small relative velocities at the time of disruption of the
constituent stars (e.g., \citealt{Jiang:2010}).
The kinematic properties of these unbound binaries and associations contain
information about the tidal field and fluctuations in the mass distribution of
the Milky Way.
With \gaia\ \DR{2}, we will be able to study the statistical properties of these
widely-separated but comoving stars at even larger separations ($>10~\pc$).

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{dx-dv.pdf}
  \end{center}
  \caption{%
    Point estimate of the 3D velocity difference between two stars in a comoving
    pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against a point estimate of the 3D
    spatial separation of the two stars, $|\bs{x}_1 - \bs{x}_2|$.
    The distances used in the point estimates are computed using a
    Lutz-Kelker-corrected parallax value (\citealt{Lutz:1973}).
    \label{fig:dx-dv}}
\end{figure}

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{separation-hist.pdf}
  \end{center}
  \caption{%
    \emph{Left}: Distribution of point estimates of the 3D spatial separation of
    the stars in all observed comoving star pairs (gray), and only those
    identified as genuinely comoving (black). Note that the contamination rate
    is $\approx 60\%$ at large separations ($>1~\pc$), but there are still a
    significant number of genuine comoving pairs at these very wide separations.
    \emph{Right}: Distribution of sky separations of all comoving star pairs
    (gray) and only those identified as genuinely comoving (black).
    \label{fig:separation}}
\end{figure}

@@

@@

\todo{Highlight the G subgiant-A pair}
\todo{Highlight a binary sequence-main sequence pair}
\todo{Highlight the largest mag difference MS pair}

\subsection{Combined with RAVE}\label{sec:rave}

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{separation-hist-rave.pdf}
  \end{center}
  \caption{%
    \emph{Left}: Distribution of point estimates of the 3D spatial separation of
    the stars in all observed comoving star pairs (gray), and only those
    identified as genuinely comoving (black). Note that the contamination rate
    is $\approx 60\%$ at large separations ($>1~\pc$), but there are still a
    significant number of genuine comoving pairs at these very wide separations.
    \emph{Right}: Distribution of sky separations of all comoving star pairs
    (gray) and only those identified as genuinely comoving (black).
    \label{fig:separation}}
\end{figure}

% \begin{figure}[htbp]
%   \begin{center}
%     \includegraphics[width=\linewidth]{genuine_cmd_rave.pdf}
%   \end{center}
%   \caption{%
%     Color-magnitude diagram of \emph{genuine} comoving star pairs plotted over
%     the density of all \tgas\ sources within 200 pc; similar to
%     \figurename~\ref{fig:sample-cmd}, but for only the genuine comoving
%     pairs.
%     Several interesting pairs that include main-sequence and giant component
%     stars are highlighted in the text
%     (\sectionname~\ref{sec:interesting-pairs}).
%     \label{fig:genuine-cmd}}
% \end{figure}

@@

Abundance of large-separation pairs

Need precise chemical information

But at very large separations, how do you convince yourself that a pair was truly coeval? Same chemistry? To what precision? What about chemical differences from planet formation?

Going forward, will have to use hierarchical mixture models to study population of singles in conjunction

Great reward: Once chemical homogeneity understood amongst stars of same spectral type, can be used to calibrate stellar models across the HR diagram: should have same age, chemistry, etc. Also tests of chemical enrichment due to planetary systems

@@

Abundances - future work, Coronado in prep.!

Dynamical capture at these radii are improbable, but <insert words about binary-binary encounters from Nathan>

@@

Low-resolution spectroscopic follow-up of comoving pairs.

Using relative RV's, confirm a sample of very wide separation comoving star pairs.

Either wide binaries, unbound binaries, or dissolving clusters

Need much larger sample with no cuts on separation or signal-to-noise to study true separation distribution

Gaia DR2 for G < 12, but really we need a concerted effort to obtain spectroscopy of wide pairs.

>>>>

\definecolor{mediumpersianblue}{rgb}{0.0, 0.4, 0.65}
\newcommand{\smoh}[1]{\textcolor{mediumpersianblue}{SO:#1}}

@@

\begin{abstract}
% Context
Widely-separated ($\gtrsim X$~pc), coeval stars are powerful tools for testing star-formation
model predictions, improving stellar atmosphere models, and inferring the
mass distribution of the Milky Way.
\smoh{The following sentence seems to be in the air while I get what you mean.}
They vastly outnumber the Galactic open clusters, but are still expected to
share a common origin and chemistry.
\smoh{How about sticking with the terminology ("comoving (star) pair") we want settled?}
The kinematic properties of comoving pairs including bound wide binaries
serve as potentiometers for the smooth tidal field and perturbation spectrum of
the Milky Way.
\smoh{If you are referring to Shaya \& Olling, how about adding 'using HIPPARCOS'?}
Thus far, a few hundred genuine comoving star pairs with separations as large
as $8~\pc$ have been discovered in the solar neighborhood.
% Aims
In previous work, we have identified $\sim 10,000$ candidate comoving star pairs
using only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS)
catalog, a part of \gaia\ \DR{1}.
Without radial velocity measurements for the vast majority of the pairs, the
\smoh{I prefer "false-positive rate" to "contamination rate" because it conveys
the meaning more directly.}
contamination rate at large separations is significant ($\sim 50\%$).
% Methods
\smoh{I think some details of telescope/instrument or just R~2000 should be mentioned here.}
Here, we present results from our own low-resolution radial-velocity survey of a sample of the
candidate pairs to validate wide-separation comoving star pairs.
% Results
Of the 311 observed comoving pairs, we identify 128 genuine comoving pairs with
separations as large as $\approx 10~\pc$, our original search limit.
% Conclusions
We find that the number of comoving star pairs with very wide separations
appears flat in linear separation at separations $\gtrsim 1~\pc$, strongly
suggesting the existence of a population of recently unbound wide binaries or
stellar associations.
% Gaia DR2 -> how many do we expect? Can we say anything? Out to what distance and
% separation for a typical MS star can we determine they are bound assuming
% high-res spectra to get very precise RVs?
\end{abstract}

@@

The majority of stars are thought to form in dynamically cold, chemically
homogeneous clusters, relative to the kinematic and chemical
dispersions observed amongst typical field stars in the Galactic disk.
Because star formation proceeds rapidly (e.g., \citealt{Bovy:2016}) and the
progenitor gas cloud disperses quickly (cite), it is typically assumed that
stars formed in the same birth cluster have the same initial chemistry
(e.g., cite + \citealt{Feng:2014}) and same ages.
These birth clusters are typically born un- or loosely-bound and therefore
quickly succumb to the tidal field of the Milky Way.
The initial kinematic associations between stars from the same cluster are then
erased as the constituent stars mix into the general population of disk stars.
However, the stars do preserve their initial associations in age, chemical
abundance patterns, and dynamical invariants (e.g., orbital actions).
It has therefore been long presumed that ``chemical tagging'' and chemodynamical
models of the Galactic disk will allow for inferences about the formation and
evolution of the Milky Way (e.g., \citealt{todo}).
That is, the chemical environment of a birth cluster is likely imprinted on its
stellar children and preserved so that their initial association may be inferred
with precisely measured abundance patterns (of $\gtrsim 10$ species).

The feasibility of chemical tagging depends on not only on the precision of
the abundance measurements but also on the intrinsic chemical
dispersions of stars formed from the same birth clusters,
and the ability to tell that the chemical abundances of
two stars from the same cluster are more similar than two randomly-paired Galactic disk stars.
Recent works suggest that, at least for red giant stars, the abundance
distributions of 20 elements in a handful of open clusters included in the
\acronym{APOGEE} survey (\citealt{Majewski:2016,Nidever:2015}) are consistent
with having intrinsic dispersions smaller than the abundance measurement
precisions ($\approx 0.03~\dex$; \citealt{Bovy:2016,Ness:2017}).
\smoh{I find the following sentence out of place. Doesn't it mean that
  the intrinsic dispersion is effectively zero rather than ``unconstrained''?}
In effect, for most elements, the intrinsic dispersions are unconstrained
because of measurement noise, so the true intrinsic dispersions are not known.
This demonstration of chemical uniformity in open clusters is promising for
chemical tagging.
However, it was also recently shown that $\approx 0.3\%$ of random pairings of
Galactic field red giant stars have abundance patterns that are
indistinguishable at the same level (``doppelgangers''), even with 20 chemical
species (\citealt{Ness:2017}).

Are these properties of open clusters universal, or limited to the few clusters
presently studied?
Do the chemical properties of open clusters depend on the star formation rate or
efficiency, which may evolve with time?
Ideal tests of the chemical homogeneity assumptions would use many more open
clusters with varied cluster masses and member stars that span a large range in
stellar masses and evolutionary stages.
Precise kinematic and color data from the \gaia\ mission
(\citealt{Gaia-Collaboration:2016,Gaia-Collaboration:2016a}) will help refine
existing membership catalogs and will lead to discoveries of many more open
clusters and young star associations (see, e.g., \citealt{Oh:2017}).
However, because they disrupt quickly (\smoh{a quantitative time scale?},
the present-day open cluster population
is biased towards younger or more massive clusters.

@@

In previous work, we have identified a sample of high-confidence, candidate
comoving star pairs (\citealt{Oh:2017}) using only astrometric information from
the \tgas\ catalog (\citealt{Michalik:2015,Gaia-Collaboration:2016a}).
These comoving pairs were selected using a cut on a marginalized likelihood
ratio computed for all nearly co-spatial pairs of stars.
Briefly, likelihood 1, $L_1$, considers a model in which the full-space
velocities of each star in the pair are identical and sampled from a disk-like,
Sun-relative velocity distribution, and likelihood 2, $L_2$, considers a model
in which the full-space velocities of each star were drawn independently from
the same velocity distribution used above. The computed likelihood ratio is
actually the ratio of fully marginalized likelihoods,
$\mathcal{L}_1/\mathcal{L}_2$, after marginalizing over true velocity and
distance for each star in a given pair.
After a parallax signal-to-noise ratio (SNR) cut ($\pi/\sigma_\pi > 8$), we
compute the likelihood ratio for all stars within $10~\pc$ of each target star
from the \tgas\ catalog.
\smoh{To be complete, we should mention $\Delta v_\perp < 10$~km/s cut.
  Or I think it's also an option to cut on much of these details consistently.}
The likelihoods appropriately take into account the reported covariances and
uncertainties between the astrometric measurements in the \tgas\ catalog and
naturally handle geometric or projection effects in comparing spherical \smoh{?}
velocity components for star pairs separated by large angles on the sky.
After applying a conservative cut on the likelihood ratio motivated by computing
the same likelihood ratio for random pairings of stars ($\ln
\mathcal{L}_1/\mathcal{L}_2 > 6$, we find 13,058 comoving star pairs with
physical separations between $10^3~\au$ and $10~\pc$.
Of these pairs, there are 10,606 unique sources, implying that a significant
fraction of the identified pairs are members of larger groups.
$\approx$4,000 of the pairs are mutually exclusive, i.e. neither member is
paired with other stars.
For a full description of the selection method, see \citealt{Oh:2017}.

@@

Spectra were obtained using the \project{Modspec} spectrograph mounted on the
$2.4~{\rm m}$ Hiltner telescope at MDM
observatory\footnote{\url{http://mdm.kpno.noao.edu}} on Kitt Peak (Arizona).
\project{Modspec} was set up in long-slit mode with a 1\arcsec\ slit, a $600~{\rm
line}~{\rm mm}^{-1}$ grating, and a $2048\times2048~{\rm pixel}$ SITe CCD
detector (``Echelle'').
Only the central 300 pixel columns were read out from the CCD, and pixel rows
$>1600$ (at the blue end) were later trimmed because of the poor quantum
efficiency of the CCD at wavelengths $\lesssim 3600~{\rm \AA}$.
The resolution and wavelength range were $\approx2~{\rm \AA}~{\rm pixel}^{-1}$
from $\approx 3600$--$7200~{\rm \AA}$.
At this resolution, to obtain high signal-to-noise spectra, most exposures were
between $30$--$120~{\rm s}$ (15th and 85th percentiles).
To maximize the number of observed pairs, we chose not to take
comparison lamp spectra at each pointing as it adds an overhead of $\approx
90~{\rm s}$ per observation.
To determine rough nightly wavelength solutions, calibration spectra were obtained using
Hg-Ne and Ne comparison lamps at the beginning and end of each night.
Additionally, atmospheric emission lines were used to correct the wavelength solution on a
per-object basis.

We observed a total of 765 unique sources over 5 nights from (UT) 2017-03-11 to
2017-03-15, corresponding to 323 unique star pairs and $\approx 120$
calibration targets.
\smoh{How about ``The calibration targets serve two purposes. First, ...
  just so that it is more explicit, plain, and direct, and also ties in
  with how we chose the calibration targets?}
The calibration targets have previously measured radial velocities and chemical
abundances, which will be used for validation of our reduction pipeline and in
future work to measure metallicities and stellar parameters for each observed
source using \project{The Cannon} (\citealt{Ness:2015}).
The spectra were reduced and calibrated using a custom, publicly-available
pipeline written in \python, described below.
\smoh{How the calibration targets were chosen is missing here.}

\smoh{Throughout this paragraph, consider changing e.g.,
  ``16 pixel rows'' to ``16 pixels in wavelength direction''?
  Or explicitly mention somewhere (perhaps at the start of this section)
  that x (along rows) is spatial and y (along columns) is spectral direction?}
For each source or comparison lamp observation, the two-dimensional image is
bias corrected, flat-field corrected, and trimmed using standard routines
implemented in \package{ccdproc} (\citealt{Craig:2015}).
For comparison lamp observations, the central 100 pixel columns are extracted
and summed along the spatial axis, weighted by the inverse-variances of each
pixel.
Source spectral traces were always placed in this region during observations,
and the curvature of the wavelength solution over this region is negligible.
For source observations, 1D spectra are extracted using an empirical
estimate of the pixel-convolved line spread function (LSF): at each of 16 pixel
rows between indices 750 and 850 (i.e. around the effective center of the
dispersion axis), a model for the LSF is fit to the 1D trace.
The LSF is assumed to be a Voigt profile with parameters $A$, amplitude; $x_0$,
profile centroid; $\sigma_G$, Gaussian root-variance; $\gamma_L$, Lorentzian
half-width at half maximum, and the background is modeled using a linear
polynomial with parameters $\alpha_1$ and $\alpha_2$, the slope and intercept
at the profile centroid, respectively.
The maximum-likelihood LSF + background model parameters are estimated from
each pixel row using \lbfgsb\ optimization (\citealt{Zhu:1994}), implemented in
the \package{scipy} package.
From the 16 fits, the median profile width parameters, $\sigma_G$ and
$\gamma_L$, are then taken to be the LSF width parameters.
At each of 1600 rows in each source spectrum image, the LSF amplitude and
centroid, and background slope and intercept at line centroid, are then fit
using the same procedure as above but with the width parameters fixed.
The LSF amplitudes are stored as the source fluxes, and the background model
intercepts are stored as the background (sky) fluxes.
Occasionally, nearby sources fall in the slit; these are masked by hand in the
2D images and do not bias the LSF extraction fits.

@@

This wavelength calibration procedure produces a nightly model for the
nonlinear dispersion of wavelength as a function of pixel value for each
extracted 1D spectrum.
However, because of flexure in the spectrograph, each observation will have an
additional offset of, typically, $\approx 0.5$--$2~{\rm \AA}$, corresponding to
a significant velocity offset at \Ha\ of $\approx 25$--$100~\kms$ (discussed
below).
\figurename~\ref{fig:sample-spectra} shows a random sample of 10
wavelength-calibrated source spectra of comoving star pair members normalized
to 1 at $\lambda = 5500~{\rm \AA}$ and \smoh{shifted by index (confusing)} for visualization.
As suggested by the features in these 10 spectra, there is a mix of
metallicities and spectral types amongst the observed targets.
Note that the spectra are not flux-calibrated, as we only care about measuring
radial velocities from this sample.

@@

Of the 323 observed star pairs, 12 have at least one member with an unreliable
(flagged) radial velocity measurement from the pipeline above.
For the remaining 311 pairs, we use the measured radial velocities to re-compute
the marginal likelihood ratio for the two hypotheses considered in previous work
(see Appendix~B, \citealt{Oh:2017}).
We compute two fully marginalized likelihoods (FMLs) for the data: (1) assuming
that the two stars are comoving with identical 3D velocity vectors (within some
allowed small dispersion), and (2) assuming that the two stars have independent
3D velocity vectors drawn from a prior velocity distribution.
As with previous work, we set the small dispersion to the orbital velocity of a
binary with the separation of the pair, and use a mixture of three Gaussian
velocity components for the prior, meant to represent the thin disk, thick disk,
and halo velocity distributions relative to the sun.

@@

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=0.7\linewidth]{llr-vs-dv.pdf}
  \end{center}
  \caption{%
    FML
    \label{fig:llr}}
\end{figure}

If the new likelihood ratio is larger than that computed from astrometry alone,
the added radial velocity information likely supports the belief that the two
stars are truly comoving in 3-space.
Conversely, if the new likelihood ratio is smaller, the radial velocity data
indicate that the pair is likely a false-positive and are not truly comoving.
In practice, we compute the logarithm of the likelihood ratios,
$\mathcal{R} = \ln \mathcal{L}_1/\mathcal{L}_2$; we refer to the
proper-motion-only log-FML-ratio for a given pair of stars as
\llrold\, and the log-FML-ratio including the radial velocity data as
\llrnew.
We consider a pair as being genuinely comoving if $\llrnew > \llrold$; 128 of
the 311 comoving pairs satisfy this constraint.
\figurename~\ref{fig:llr} shows the magnitude of the 3D velocity difference
plotted against the change in log-likelihood-ratio for the two stars in all
observed pairs.
In detail, for visualization, we sample from the error distributions over proper
motion, parallax, and radial velocity for each star individually and compute the
magnitude of the 3D velocity differences, $|\bs{v}_1-\bs{v}_2|$, for $2^{16}$
samples.
We plot the median of the resulting distribution over velocity-difference as
dark markers (circles).
The error bars show the 15th and 85th percentiles of the distributions for a few
randomly chosen pairs.
Note that pairs with large 3D velocity differences have large (negative) changes
to their log-likelihood-ratio values.

@@

\todo{Discuss contamination rate vs. separation for ratio cut of 6?}
At separations $\gtrsim 1~\pc$, $\approx 40\%$ of the comoving pairs are
genuinely comoving; interestingly, the separation distribution of the genuine
pairs appears to be at least uniform or possibly growing in number as the
separation increases.
This strongly suggests the existence of a population of unbound wide binaries or
dissolved open clusters, which maintain approximate spatial coherence for long
times because of the small relative velocities at the time of disruption of the
constituent stars (e.g., \citealt{Jiang:2010}).
The kinematic properties of these unbound binaries and associations contain
information about the tidal field and fluctuations in the mass distribution of
the Milky Way.
With \gaia\ \DR{2}, we will be able to study the statistical properties of these
widely-separated but comoving stars at even larger separations ($>10~\pc$).

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{dx-dv.pdf}
  \end{center}
  \caption{%
    FML
    \label{fig:dx-dv}}
\end{figure}

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{separation-hist.pdf}
  \end{center}
  \caption{%
    FML
    \label{fig:separation}}
\end{figure}

@@

\todo{separation distribution - compare with Shaya and Olling sample!}
\todo{10 overlap with Shaya and Olling}
\todo{RAVE pairs - both observed}

@@

\todo{Highlight the MS-Giant pairs}
\todo{Highlight the binary sequence-main sequence pairs}

@@

Once chemical homogeneity understood amongst stars of same spectral type, can be used to calibrate stellar models across the HR diagram: should have same age, chemistry, etc. Also tests of chemical enrichment due to planetary systems

@@

Abundances!

- Stellar rotation of evolved-main sequence
- Full catalog should contain Shaya and Olling, cross-match to RV surveys

@@
